Scaling of Electrode-Electrolyte Interface Model Parameters In Phosphate Buffered Saline

### 2013-11-14 paper.tex
Modified abstract to be over 100 words (now 105). Added information about animal ethics. Reworded part abouot testing PBS as substitute for CSF (as advised by Peter). Changed paper format to match that of Jonathans submission to TBioCAS. Changed blurb about myself to match that of similar transactions. Added acknowledgements for UOW and Saluda.

\documentclass[10pt,final,journal]{IEEEtran}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{url}

\title{Scaling of Electrode-Electrolyte Interface Model Parameters In Phosphate Buffered Saline}
\author{Mark~H.~Jones,~Jonathan~Scott,~\IEEEmembership{Senior Member,~IEEE}}
\begin{document}
\maketitle

\begin{abstract}
We report how the impedance presented by a platinum electrode scales with the concentration of phosphate buffered saline (PBS). 
We find that the constant phase element of the model scales with approximately the log of concentration, whereas the resistivity is inversely proportional. 
Using a novel DC measurement technique we show that the Faradaic response of a platinum electrode, and thus the safe exposure limit, does not scale with concentration below 900\thinspace mV overpotential across a pair of electrodes. 
We comment upon the appropriateness of PBS as a substitute for the spinal cavity of a sheep.
\end{abstract}

\section{Introduction}
There is considerable interest in the electrical modelling of electrodes. \cite{Cogan2008,Troy2006}
In \cite{Franks2005} a linearised model was presented, and in \cite{ScottSingle2013} a non-linear model
suitable for use with the SPICE circuit simulator was presented. The SPICE model usefully characterises an electrode in a given electrolyte with a small number of parameters.
One major reason for the interest in the electrical impedance of electrodes concerns the design of electronics intended for integration into pacemakers, cochlear implants, spinal cord stimulators, etc. The design of successful circuits depends upon a good understanding of external load impedance, while maximisation of battery life is linked to the use of electrodes whose impedance is well understood.
Having a compact model of an electrode in the appropriate electrolyte allows circuit designers to simulate their designs with valid loads. Knowing the effect upon the model of changes in the electrode or electrolyte would allow designers to anticipate circumstances that will affect the load seen by their circuits. Another appeal of a compact model stems from the fact that electrode characteristics are succinctly and objectively represented by the model parameters. These enable direct comparison of different electrodes, or the change in electrode properties over time. For example, in \cite{Kane13} changes in chronically-implanted electrodes are observed but there is no standard quantitative way of presenting the changes. Similarly, while there is a good understanding of what represents safe exposure, a compact model with parameters permits the prediction of the safety of any given stimulus regime.~\cite{Merrill05}

@@

\begin{figure}
    \begin{center}
        \includegraphics[width=230pt]{graphics/interfaceSchematic_noMemristive.eps}
    \end{center}
    \caption{Electrical schematic of the electrode-interface impedance model used in this work. $D_{a}$ and $D_{b}$ represent diodes; $CPE$ and $R_{s}$ represent the constant phase element and series resistance respectively.}
    \label{fig:schematic}
\end{figure}

@@

\begin{figure}
    \begin{center}
        \includegraphics{graphics/pbs_transimpedance_IEEE}
    \end{center}
    \caption{Trans-resistance measurements (markers) used to generate the resistor mesh and the corresponding fit (lines). Red diamonds indicate results where a stimulus is placed across electrodes 1 and 2. Blue circles represent measurements where the stimulus is placed across electrodes 1 and 8. Each trace represents one of six concentrations of PBS, increasing in transresistance magnitude monotonically from 1.0X to 0.025X.}
    \label{fig:transresistance}
\end{figure}

@@

We now test the claim that 0.1X PBS is a reasonable representation of cerebral spinal fluid (CSF).
The displacement measurements were repeated in a live sheep. The sheep was prepared and anaesthetised using the procedures described in \cite{Parker2013}. The measurements are presented in figures~\ref{fig:displacement_sheepCPEMagnitude} and~\ref{fig:displacement_sheepCPEPhase} on top of the array of traces for saline taken from figures~\ref{fig:CPE_Magnitude}
and~\ref{fig:CPE_Phase}. 
Although it is clear that there are complications in vivo, a 0.25X PBS saline offers a good approximation to the resistive part of the trace.
The reactive part would be better modelled by a solution of less than 0.025X concentration.

@@

@@

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{graphics/MarkJones-Aug13.jpg}}]{Mark Jones}
graduated with a BSc in Physics and a Masters in Microwave Electronics from the University of Waikato in 2008 and 2009.
He is currently studying towards his PhD.
\end{IEEEbiography}

>>>>

\documentclass[journal, a4paper]{IEEEtran}

\usepackage{cite}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{url}

\begin{document}

\title{Scaling of Electrode-Electrolyte Interface Model Parameters In Phosphate Buffered Saline}

\author{Mark~H.~Jones,~Jonathan~Scott,~\IEEEmembership{Senior Member,~IEEE}

\thanks{Mark~H.~Jones is with the School of Electronic Engineering, University of Waikato, New Zealand, e-mail: markjones112358@gmail.com.}%

\thanks{Jonathan Scott is with the School of Electronic Engineering, University of Waikato, New Zealand, e-mail: jonathanscott@iee.org.}

}

\markboth{Transactions on Biomedical Circuits and Systems}
{Jones \MakeLowercase{\textit{et al.}}: Scaling of Electrode-Electrolyte Interface Model Parameters in Phosphate Buffered Saline\...}
\maketitle

\begin{abstract}
We report how the impedance presented by a platinum electrode scales with the concentration of phosphate buffered saline (PBS).
We find that the constant phase element of the model scales with approximately the log of concentration, whereas the resistivity is inversely proportional.
Using a novel DC measurement technique we show that the Faradaic response of a platinum electrode, and thus the safe exposure limit, does not scale with concentration below 900\thinspace mV overpotential across a pair of electrodes.
We compare objective measurements made in saline to those made in the spinal cavity of live sheep. We comment upon the appropriateness of using PBS as a substitute for living sheep.
\end{abstract}

\begin{IEEEkeywords}
    Electrical stimulation, Bioelectric phenomena, Biophysics, Bioimpedance, Biomedical electrodes, Biomedical measurements, Implantable biomedical devices
\end{IEEEkeywords}

\section{Introduction}
\IEEEPARstart{T}{here} is considerable interest in the electrical modelling of electrodes. \cite{Cogan2008,Troy2006}
In \cite{Franks2005} a linearised model was presented, and in \cite{ScottSingle2013} a non-linear model
suitable for use with the SPICE circuit simulator was presented. The SPICE model usefully characterises an electrode in a given electrolyte with a small number of parameters.
One major reason for the interest in the electrical impedance of electrodes concerns the design of electronics intended for integration into pacemakers, cochlear implants, spinal cord stimulators, etc. The design of successful circuits depends upon a good understanding of external load impedance, while maximisation of battery life is linked to the use of electrodes whose impedance is well understood.
Having a compact model of an electrode in the appropriate electrolyte allows circuit designers to simulate their designs with valid loads. Knowing the effect upon the model of changes in the electrode or electrolyte would allow designers to anticipate circumstances that will affect the load seen by their circuits. Another appeal of a compact model stems from the fact that electrode characteristics are succinctly and objectively represented by the model parameters. These enable direct comparison of different electrodes, or the change in electrode properties over time. For example, in \cite{Kane13} changes in chronically-implanted electrodes are observed but there is no standard quantitative way of presenting the changes. Similarly, while there is a good understanding of what represents safe exposure, a compact model with parameters permits the prediction of the safety of any given stimulus regime.~\cite{Merrill05}

@@

\begin{figure}
    \begin{center}
        \includegraphics[width=230pt]{graphics/interfaceSchematic_noMemristive}
    \end{center}
    \caption{Electrical schematic of the electrode-interface impedance model used in this work. $D_{a}$ and $D_{b}$ represent diodes; $CPE$ and $R_{s}$ represent the constant phase element and series resistance respectively.}
    \label{fig:schematic}
\end{figure}

@@

\begin{figure}
    \begin{center}
        \includegraphics{graphics/pbs_transimpedance_IEEE}
    \end{center}
    \caption{Transresistance measurements (markers) used to generate the resistor mesh and the corresponding fit (lines). Red diamonds indicate results where a stimulus is placed across electrodes 1 and 2. Blue circles represent measurements where the stimulus is placed across electrodes 1 and 8. Each trace represents one of six concentrations of PBS, increasing in transresistance magnitude monotonically from 1.0X to 0.025X.}
    \label{fig:transresistance}
\end{figure}

@@

We now test the claim that 0.1X PBS is a reasonable representation of the spine of a living sheep.
The displacement measurements were repeated inside the spinal canal of a live sheep, just outside the dura containing CSF. The sheep were prepared and anaesthetised using the procedures described in \cite{Parker2013} under the Animal Care and Ethics Committee approval of the Royal North Shore Hospital. The study complied with the Australian Code of Practice for the Care and Use of Animals for Scientific Purposes. The measurements are presented in figures~\ref{fig:displacement_sheepCPEMagnitude} and~\ref{fig:displacement_sheepCPEPhase} on top of the array of traces for saline taken from figures~\ref{fig:CPE_Magnitude}
and~\ref{fig:CPE_Phase}.
Although it is clear that there are complications in vivo, a 0.25X PBS saline offers a good approximation to the resistive part of the trace.
The reactive part would be better modelled by a solution of less than 0.025X concentration.

@@

\section*{Acknowledgment}
The authors wish to acknowledge the The University of Waikato for financial support and Saluda Medical for the use of facilities and equipment.

@@

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{graphics/MarkJones-Aug13.jpg}}]{Mark Jones}
received the B.Sc. degree in Physics in 2008 and the M.E. degree in Microwave Electronics in 2009 from The University of Waikato, Hamilton, New Zealand.
He is working towards a Ph.D. in Electrical Engineering, also at The University of Waikato.
\end{IEEEbiography}
